\chapter{A Final Word}

One of the sad results of \emph{charismania} is its repelling influence on so many hungry, earnest saints of God who are needing and searching for a deeper experience of God's power in their lives. 

We read in II Kings 4 that Elisha, during a famine, sought to feed the sons of the prophets by setting before them a great seething pot into which they had placed the herbs that they had gathered. One of the young prophets had shredded wild gourds into the pottage so that when they sought to satisfy their hunger by eating they cried out, "There is death in the pot!" 

This is what often happens when a person, hungry for the fullness of God, tastes of the noxious wildfire of the \emph{charismaniacs}. They unfortunately conclude that there is no genuine work of God's Spirit in the church today, and they continue to struggle with their Christian walk without the help of the full dynamic of the Holy Spirit. 

God did not intend the children of Israel to die in the wilderness, but to come into the overflowing abundance of the Promised Land. God does not intend that your walk with Him be a dry, barren wilderness experience, but He desires that you come to know the full richness of that life promised to us through the Holy Spirit. 

Do not allow the unscriptural excesses of those practicing \emph{charismania} to discourage you from seeking all that God wants you to experience of the love, joy, and power of living in the fullness of the Spirit. We have not yet experienced the full, rich spectrum of the genuine work of the Spirit in our lives; thus we need to always remain open to God for whatever He may desire to bestow upon us. There is so much in the Scriptures which we have not yet experienced that we certainly have no need to go beyond the Scriptures. 

Paul expressed his joy that the Corinthian church did not lack or come short of any spiritual gift as they awaited the coming of Jesus. So my prayer is that we also may experience all of the fullness of the Holy Spirit and the gifts that He desires to bestow upon us as we wait for our Lord's return. 


